---
title: "preprocessing"
author: "Laura.w"
date: "12/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Read csv file into a dataframe
install.packages("readr")
library(readr)
install.packages("tibble")
library(tibble)
install.packages("dplyr")
library(dplyr)

```


################################### Get the Data Ready ############################

```{r}
#load data

wb <- read_csv("C:/Users/LW/Documents/R/LE_FV/wb.csv")

typeof(wb)
is.data.frame(wb)
is_tibble(wb)

#view the data
View(wb)
head(wb)


#obtain a summary of the data: data types of the columns
#str(wb)
summary(wb)
```
```{r}


```

There's an average of 47.35 rate of women who have a decision in their reproductive health.
```{r}

wb$womenrephealth <- as.numeric(wb$`Women making their own informed decisions regarding sexual relations, contraceptive use and reproductive health care  (% of women age 15-49) [SG.DMK.SRCR.FN.ZS]`)
  
mean(wb$womenrephealth, na.rm = TRUE)
median(wb$womenrephealth, na.rm= TRUE)
sort(table(wb$womenrephealth))

#5-number summary
summary(wb$womenrephealth)
var(wb$womenrephealth, na.rm =TRUE)
```

######### Plotting #################

```{r}
#see some skewness (skew right or positively skewed, meaning more younger than older people are victims)
boxplot(wb$womenrephealth)

```

#check the normality of age distribution using QQ normal plot: 
#if data were normaly distributed, the age should have been higher than observed

```{r}
qqnorm(wb$womenrephealth)
```

```{r}
#histogram
hist(wb$womenrephealth)
```

##################### Take a look at other attributes #################
```{r}
#want to know unique values in some of the columns

wb$womenfertrate <- as.numeric(wb$`Fertility rate, total (births per woman) [SP.DYN.TFRT.IN]`)

wb$pop <- as.numeric(wb$`Population ages 15-64, total [SP.POP.1564.TO]`)

wb$country <- as.factor(wb$`Country Name`)


unique(wb$womenfertrate)
unique(wb$country)
#added groups of countries

```

```{r}
tail(wb)
wb$LE <- as.numeric(wb$`Life expectancy at birth, total (years) [SP.DYN.LE00.IN]`)

unique(wb$LE)

#distribution in 
table(wb$LE)/sum(table(wb$LE)) #2/3 cases not fleeing

boxplot(wb$LE)

```
```{r}
#get a sense of the extend of missing data (lots of values for race is missing)
#815 out of 4478 rows has some missing value
filter(wb, !complete.cases(wb))

#apply() returns a vector or array or list of values obtained by applying a function to margins of an array or matrix.
#number of NAs in each column ('2')
apply(wb, 2, function(x) sum(is.na(x))) %>% sort(decreasing=TRUE) 
#What shall we do with NAs? Not clear at this time, so keep NAs as is. 

#which stae/city has most shooting cases?
sort(table(wb$pop), decreasing =FALSE)
dplyr::count(wb, pop, sort = TRUE)
summary(wb$pop)

```

################## STATES ####################

```{r}
install.packages("ggplot2")
library(ggplot2)


```


```{r}
wb$LE <- as.factor(wb$`Life expectancy at birth, total (years) [SP.DYN.LE00.IN]`)

#Life Expectancy distribution by race
ggplot(wb, aes(x=womenrephealth)) +
  geom_bar() +
  ggtitle("reproduction health choices")

```

```{r}
#shooting distribution by states
ggplot(wb, aes(x=LE, fill=)) +
  geom_bar() +
  ggtitle("shooting distribution by states")

```




#bring the race factor in
ggplot(shooting_orig, aes(x=state, fill=race)) +
  geom_bar() +
  ggtitle("shooting distribution by states")

#now I see the need to make the x-tick marks more readable
#http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels
ggplot(shooting_orig, aes(x=state, fill=race)) +
  geom_bar() +
  ggtitle("shooting distribution by states") +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(size=6, angle=90), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

#Are the number of incidence correlated with state population?
#Bring popluation data in to the picture

install.packages("usmap")
library(usmap)

#usmap data has regions and poverty info we could use
statepop #state population
statepov #state poverty index

#is there a correlation between incident count and population or poverty level of the states?
#find incident count for each of the states
(state_count <- arrange(count(shooting_orig, state), desc(n)))

#create joint dataframe by matching on state abbreviations to run correlation

#need to join on 'state', so change colname 'abbr' used in statepop/statepov to 'state'
colnames(statepop)[colnames(statepop) == "abbr"] <- "state"
colnames(statepov)[colnames(statepov) == "abbr"] <- "state"
#use dplyer inner_join to join state incident count and state population/poverty tables
(state_join_pop <- inner_join(state_count, statepop[, c("state", "pop_2015")], by=c("state")))
(state_join_pov <- inner_join(state_count, statepov[, c("state", "pct_pov_2014")], by=c("state")))

#correlation btw shooting and population size and poverty level
#scatter plot
plot(state_join_pop$n, state_join_pop$pop_2015, main="Relationships btw shooting count and population size")
#correlation coefficience
cor(state_join_pop$n, state_join_pop$pop_2015)
#population size is highly correlated

plot(state_join_pop$n, state_join_pov$pct_pov_2014, main="Relationships btw shooting count and porverty level")
cor(state_join_pop$n, state_join_pov$pct_pov_2014)
#poverty level is not correlated

View(arrange(statepov, desc(pct_pov_2014)))


############### relationship with date?  => year, month
install.packages("lubridate")
library(lubridate)

shooting_orig <- mutate(shooting_orig,
  year = year(shooting_orig$date),
  month = month(shooting_orig$date)
)

colnames(shooting_orig)
shooting_orig[, c('date', 'year', 'month')]

ggplot(shooting_orig, aes(x=month)) +
  geom_bar() +
  ggtitle("shooting distribution by month")
#winter months have fewer shootings, why? Sept - Dec

ggplot(shooting_orig, aes(x=month)) +
  geom_bar() + facet_grid("year~race")
  ggtitle("shooting distribution by month")
